On the presentation of orthogonal groups over polynomial rings

### 2019-06-17 dl-paper.tex
Moved lemmas

\begin{comment}
Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$ 
\end{comment}
The classical Matsumoto theorem (see~\cite[Theorem~5.10]{Ma69}) allows one to compute the group $\K_2(\Phi, R)$ in the special case when $R=k$ is a field. More concretely, it asserts that $\K_2(\Phi, k) \cong \K_2^\mathrm{MW}(k)$ provided $\Phi$ is symplectic and  $\K_2(\Phi, k) \cong \K_2^\mathrm{M}(k)$ otherwise. In the following lemma we compute the group $\K_2(\Phi, R)$ in yet another case when $R=k[X, X\inv]$.
\begin{lemma}\label{K2-laurent-field} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be arbitrary field. Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ in which} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ k^* & \text{otherwise.}  \end{array}\right. \]  \end{lemma}
\begin{proof} If $\Phi$ is nonsymplectic one can choose a long simple root $\alpha\in \Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^* \ar[r]^-{\{ -, X \}_{\alpha}} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)).} \]
Clearly, the diagonal arrow is injective by Milnor's theorem and the vertical arrow is an isomorphism by Matsumoto theorem. The assertion now follow from~\cite[Satz~3]{Hur77} and the fact that the left horizontal arrow is injective.

If $\Phi$ is symplectic the assertion follows from~\cite[Theorem~B]{MR91}, provided one verifies that Morita--Rehmann's group $P(F)$ is isomorphic to $\K_1^\mathrm{MW}(F)$,
 which is an easy exercise. Alternatively, this isomorphism can be obtained as a corollary of~\cref{bass-ft} and~\cite[Lemma~4.1.1]{AF17}. \end{proof}

@@

\section{First injectivity theorem}
\begin{corollary} \label{field-injectivity} For $\Phi\neq\rG_2$ the map $\St(\Phi, k[X]) \to \St(\Phi, k[X, X^{-1}])$ is injective. \end{corollary}
\begin{proof} The result follows from consideration of the commutative diagram
\[\xymatrix{ & \K_2(\Phi, k) \ar[dl]_{\cong} \ar@{^{(}->}[dr] & \\
               \K_2(\Phi, k[X]) \ar[rr] &               & \K_2(\Phi, k[X, X^{-1}])} \]
in which the left arrow is an isomorphism by the corollary of~\cite[Satz~1]{Re75} and the right arrow is split injective by~\cref{K2-laurent-field}. \end{proof}
\begin{theorem} \label{thm41}
Assume the notation of~\cref{lem:lemma32} and suppose additionally that $2$ is invertible in $A$.
Then for $\ell \geq 7$ the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}
 Consider the following commutative diagram with exact rows, in which the lifting $t$ is obtained from~\cref{lem:lemma32}:
\begin{equation*} \xymatrix{
 C(\rD_\ell, B, I) \ar[r]^{j_B} \ar@{->>}[d]_{f} & \St(\rD_\ell, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\rD_\ell, B) \ar[r]^{\pi_B} \ar[d]_{h} & \St(\rD_\ell, k[X]) \ar[d]_{i} \\
 C(\rD_\ell, R, I) \ar[r]^{j_R}         & \St(\rD_\ell, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\rD_\ell, R) \ar[r]^-{\pi_R}        & \St(\rD_\ell, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$
 also lies in $\Ker(\pi_B)$ and hence comes from some $b \in \St(\rD_\ell, B, I)$ via $\mu_B$.
Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\rD_\ell, R, I)$ such that $j_R(c) = g(b)$. 
By~\cref{lem:prop41}, $f$ is surjective, therefore $c = f(d)$ for some $d \in C(\rD_\ell, R, I)$ and we obtain the required assertion:
 \[ 1 = \mu_Bj_B(d) = tgj_B(d) = tj_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]
\end{proof}

\begin{comment}
\[x_{\alpha}(aX^{-1}) x_{-\alpha}(mX) x_{\alpha}(-a(1+am)^{-1}X^{-1}) \cdot \{X, 1+am\} \in G_+.\]
$y := x_{\alpha}(aX^{-1}) x_{-\alpha}(mX) x_{\alpha}(-a(1+am)^{-1}X^{-1})$, 
$z = x_{-\alpha}(-m(1+am)^{-1}X) y h_{\alpha}((1+am)^{-1})$,
Using $\pi(z) = 1$ we get 
\begin{multline}z = z^{h_{ik}^{-1}(X)} = \\ x_{ji}(-m(1+am)^{-1}) x_{ij}(a) x_{ji}(m) \cdot x_{ij}(-a(1+am)^{-1})\{X, (1+am)^{-1}\} \cdot h_{ij}((1+am)^{-1}) \in G. \end{multline}

We need to introduce notation for certain subgroups of $\St(A[X, X^{-1}])$:
\begin{align}
 G_+^0 & = \mathrm{Im}(\St(\Phi, A[X], XM[X]) \to \St(A[X, X^{-1}]))\\
 G_+   & = \mathrm{Im}(\St(\Phi, A[X], M[X]) \to \St(A[X, X^{-1}]))
\end{align}
Notice that for $ g\in G_+$ the element $ev_{X=0}(g)^{-1}g$ lies in $G_+^0$.

\begin{rem} Our goal is to rewrite $x_\alpha(aX^{-1}) \cdot z_\beta(f_1(X), f_2(X))$, where $f_1(X) \in M[X]$, $f_2(X) \in A[X]$ as
 $g \cdot x_\alpha(\frac{aX^{-1}}{1+am}) \cdot \{ 1 + am, X \}$ for some $g \in G_+$
 and $m \in M$. At first assume that $\Phi$ is simply-laced. \end{rem}

\begin{lemma} \label{lem:lem34} Assume that $\Phi$ is simply-laced and of rank $\geq 2$.
 For any $a\in A$, $\alpha\in \Phi$, $g \in G^0_+$ there exists $m = m(g)\in M$ and $g'\in G_+$ such that 
 \[x_{\alpha}(aX^{-1}) g  = g' Y_\alpha(a,m) \text{ where } Y_\alpha(a,m) = x_\alpha(-a(1+am)^{-1}X^{-1}) \cdot \{X, 1+am\}.\]
\end{lemma}
\begin{proof}
 We can choose some parabolic subset $S$ of roots in $\Phi$ in such a way that its special part $\Sigma_S$ does not contain the root $\alpha$.
 By~\cref{lem:Zgen} the subgroup $G_+^0$ is generated by images in $\St(\Phi, A[X, X^{-1}]$ of the elements of the set $\mathcal{Z}(\Sigma_S, A[X], XM[X])$.
 Thus, we can write $g_0$ as a product of elements $z_\beta(f_1(X), f_2(X))$ such that $f_1(X) \in XM[X]$, $f_2(X) \in A[X]$ and, moreover, $f_2(X) = 0$ whenever $\beta \not \in \Sigma_S$.

 It is easy to verify using~\eqref{eq:symbol-properties} that \[x_\alpha(aX^{-1}) \cdot g_0 \cdot g_1 = g_0' \cdot Y_\alpha(a, m(g_0)) \cdot g_1 = g_0' \cdot g_1' \cdot Y_\alpha(a, m(g_0)+m(g_1)),\] therefore it suffices demonstrate the assertion of the lemma in the special case $g = z_\beta(f_1(X), f_2(X))$.

 We proceed by case analysis. Let us first analyze the cases when $\beta \neq -\alpha$. In all these cases $m(g) = 0$.
 \begin{enumerate}
 \item {\it Case $\alpha \pm \beta \not \in \Phi \cup \{ 0\}$.}  Obviously, in this case also $g' = g$.
 \item {\it Case $\alpha + \beta \in \Phi$, $f_1(X) = Xf_1'(X)$, $f_2(X) = 0$.}
 \[ x_\alpha(aX^{-1}) \cdot x_\beta(Xf_1'(X)) = x_{\alpha+\beta}(af_1'(X))x_\beta(Xf_1'(X)) x_\alpha(aX^{-1}) \]
 \item {\it Case $\alpha = \beta$.} By our choice of $S$ holds $\beta \not \in \Sigma_S$, therefore $f_2(X) = 0$ and $g$ commutes with $x_\alpha(aX^{-1})$.
 \item {\it Case $\alpha + \beta \in \Phi$, $f_1(X) = Xf_1'(X)$.}
 Direct computation using formula from~\cite[Lemma~9]{S15} shows that:
 \begin{multline*}
 x_\alpha(aX^{-1}) \cdot z_\beta(f_1(X), f_2(X)) = \\ = x_\alpha(af_1'(X)f_2(X)) \cdot x_{\alpha+\beta}(-N_{\beta, \alpha} af_1'(X)) \cdot z_\beta(f_1(X), f_2(X)) \cdot x_\alpha(aX^{-1}); 
 \end{multline*}
 \item {\it Case $\alpha - \beta \in \Phi$, $f_1(X) = Xf_1'(X)$.} Similarly to the above case:
 \begin{multline*}
  x_{\alpha}(aX^{-1}) \cdot z_\beta(f_1(X), f_2(X)) = \\ = x_{\alpha}(-af_1'(X)f_2(X))\cdot x_{\alpha-\beta}(-N_{\beta,-\alpha} af_1'(X)f_2(X)^2) \cdot z_\beta(f_1(X), f_2(X)) \cdot x_{\alpha}(aX^{-1}).
 \end{multline*} \end{enumerate}
Now we turn to the case $\beta = -\alpha$, $f_1(X) = Xf_1'(X)$.
Since $g$ commutes with $x_\alpha(\pm f_2(X))$, it suffices to prove the assertion in the special case $f_2(X)=0$.
First consider the case $f_1(X) = mX$, where $m\in M$. Since $\Phi$ is nonsymplectic, we can choose $\gamma$ such that $\langle \alpha, \gamma \rangle = -1$.
Now direct computation invoking~\eqref{eq:conj-h-h}--\eqref{eq:dennis-stein} and the centrality of symbols shows that:
\begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(X m) %= x_\alpha(X^{\langle \alpha, \gamma \rangle}a) x_{-\alpha}(X^{-\langle \alpha, \gamma \rangle}m) 
 = {}^{h_\gamma(X)}(x_\alpha(a) x_{-\alpha}(m)) = \\
 = {}^{h_\gamma(X)}\left( x_{-\alpha}\left(\frac{m}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_\alpha\left(\frac{a}{1+am}\right) \right) = \\
 x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(X^{-1}(1+am))\cdot h_\alpha(X^{-1})^{-1} \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot \{1+am, X^{-1}\}^{-1} \cdot h_\alpha(1+am)\cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) = \\
 = x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \cdot h_\alpha(1+am) \cdot x_{\alpha}\left(\frac{aX^{-1}}{1+am}\right) \cdot \{1+am, X\}.
\end{multline*}

Now suppose $f_1(X) = X^2f_1'(X)$. Choose $\gamma$ such that $\alpha + \gamma \in \Phi$.
Then everything we need follows from the computation:
\begin{multline*}
 x_\alpha(aX^{-1}) \cdot x_{-\alpha}(X^2 mf_1'(X)) \cdot x_\alpha(-aX^{-1}) = 
 {}^{x_\alpha(aX^{-1})}[x_{-\alpha - \gamma}(N_{-\alpha-\gamma, \gamma}Xf_1'(X)), x_{\gamma}(X)] = \\
 = [x_{-\alpha-\gamma}(N_{-\alpha-\gamma, \gamma}Xf_1'(X)) x_{-\gamma}(N_{-\alpha-\gamma, \gamma} N_{-\alpha-\gamma, \alpha} af_1'(X)), x_\gamma(X) x_{\alpha+\gamma}(N_{\alpha,\gamma}a)].
\end{multline*} 
\end{proof}

\begin{rem}
 From the proof of the above lemma it is easy to describe the way to compute... 
\end{rem}
\include{rehmann}
\end{comment}

\section{Second injectivity theorem}
\begin{comment} 
First of all, set $\widetilde{V} = \overline{G}^{\geq 0}_M \times G^{\leq 0} \times (1 + M)^*$
 and consider it as a pointed set, i.\,e. ignore its structure of the group but retain $\widetilde{v_0} = (1, 1, 1)$ as a choice of a basepoint.
We let the group $\St(\Phi, R, M)$ act on $\widetilde{V}$ via the formula 
 $h \cdot (\overline{g}^+, g^-, u) = (\overline{g}^+ \cdot j_+i_+(h), i_-(h)^{-1}\cdot g^-, u)$.
Denote by $V$ the set of orbits of this action with the fixed basepoint $v_0 = O(\widetilde{v}_0)$.
We will use the notation $[\overline{g}^+, g^-, u]$ as a shorthand for $O(\overline{g}^+, g^-, u)$.

It is obvious that the map $\widetilde{p} \colon \widetilde{V} \to \St(\Phi, R[t, t^{-1}])$ given by 
 \[(\overline{g}^+, g^-, u) \mapsto \overline{g}^+ \cdot j_-(g^-) \cdot \{t, u\}\]
 is constant on the orbits of the above action
  and therefore gives rise to a well-defined map $p \colon V \to \St(\Phi, R[t, t^{-1}])$ of pointed sets.

We will construct an action of $\St(\Phi, B)$ on the set $V$ with the following additional properties:
\begin{enumerate}
\item for $g^- \in \St(\Phi, R[t^{-1}])$ one has $\varphi(g^-) \cdot v_0 = [1, g^-, 1]$, where $\varphi$ is the canonical map
  $\St(\Phi, R[t^{-1}]) \to \St(\Phi, B)$ induced by the embedding $R[t^{-1}] \to B$.
\item for $g \in \St(\Phi, B)$ the element $p(g \cdot v_0)$ coincides with the canonical image of $g$ in $\St(\Phi, R[t, t^{-1}])$; 
\end{enumerate} \end{comment}

>>>>

\begin{comment}
Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$ 
\end{comment}
The classical Matsumoto theorem (see~\cite[Theorem~5.10]{Ma69}) allows one to compute the group $\K_2(\Phi, R)$ in the special case when $R=k$ is a field. More concretely, it asserts that $\K_2(\Phi, k) \cong \K_2^\mathrm{MW}(k)$ provided $\Phi$ is symplectic and  $\K_2(\Phi, k) \cong \K_2^\mathrm{M}(k)$ otherwise. In the following lemma we compute the group $\K_2(\Phi, R)$ in yet another case when $R=k[X, X\inv]$.
\begin{lemma}\label{K2-laurent-field} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be arbitrary field. Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ in which} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ k^* & \text{otherwise.}  \end{array}\right. \]  \end{lemma}
\begin{proof} Let us first consider the case of nonsymplectic $\Phi$, in which one can find a long root $\alpha\in \Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^* \ar[r]^-{\{ -, X \}_{\alpha}} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)).} \]
Clearly, the diagonal arrow is injective by Milnor's theorem and the vertical arrow is an isomorphism by Matsumoto theorem. The assertion now follow from~\cite[Satz~3]{Hur77} and the fact that the left horizontal arrow is injective.

In the symplectic case the assertion follows from~\cite[Theorem~B]{MR91}, provided one verifies that Morita--Rehmann's group $P(F)$ is isomorphic to $\K_1^\mathrm{MW}(F)$,
 which is an easy exercise. Alternatively, this isomorphism can be obtained as a corollary of and~\cite[Lemma~4.1.1]{AF17} and \cref{bass-ft} below. \end{proof}

\begin{corollary} \label{field-injectivity} For $\Phi\neq\rG_2$ the map $\St(\Phi, k[X]) \to \St(\Phi, k[X, X^{-1}])$ is injective. \end{corollary}
\begin{proof} The result follows from consideration of the commutative diagram
\[\xymatrix{ & \K_2(\Phi, k) \ar[dl]_{\cong} \ar@{^{(}->}[dr] & \\
               \K_2(\Phi, k[X]) \ar[rr] &               & \K_2(\Phi, k[X, X^{-1}])} \]
in which the left arrow is an isomorphism by the corollary of~\cite[Satz~1]{Re75} and the right arrow is split injective by~\cref{K2-laurent-field}. \end{proof}

@@

\section{First injectivity theorem}

\begin{theorem} \label{thm41}
Assume the notation of~\cref{lem:lemma32} and suppose additionally that $2$ is invertible in $A$.
Then for $\ell \geq 7$ the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}
 Consider the following commutative diagram with exact rows, in which the lifting $t$ is obtained from~\cref{lem:lemma32}:
\begin{equation*} \xymatrix{
 C(\rD_\ell, B, I) \ar[r]^{j_B} \ar@{->>}[d]_{f} & \St(\rD_\ell, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\rD_\ell, B) \ar[r]^{\pi_B} \ar[d]_{h} & \St(\rD_\ell, k[X]) \ar[d]_{i} \\
 C(\rD_\ell, R, I) \ar[r]^{j_R}         & \St(\rD_\ell, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\rD_\ell, R) \ar[r]^-{\pi_R}        & \St(\rD_\ell, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$
 also lies in $\Ker(\pi_B)$ and hence comes from some $b \in \St(\rD_\ell, B, I)$ via $\mu_B$.
Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\rD_\ell, R, I)$ such that $j_R(c) = g(b)$. 
By~\cref{lem:prop41}, $f$ is surjective, therefore $c = f(d)$ for some $d \in C(\rD_\ell, R, I)$ and we obtain the required assertion:
 \[ 1 = \mu_Bj_B(d) = tgj_B(d) = tj_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]
\end{proof}

\section{Second injectivity theorem}
